---
title: "MAPP_run_therMizer_ResourceCalibration"
author: "Phoebe.Woodworth-Jefcoats@noaa.gov"
date: "2024-01-23"
output: html_document
---

## Model runs

This script runs _therMizer_ for calibrating the plankton spectrum that's informed by the LENS output.  

Set up the environment:
```{r}
# Load libraries
library(mizer)
library(here)
library("remotes")
remotes::install_github("sizespectrum/therMizer")
library("therMizer")
```

Set up the species parameters:
```{r}
# Load parameters 
HIparams <- read.csv("HIregion_species_params.csv", header = TRUE)
HIinter <- read.csv("HIregion_inter.csv", row.names = 1)
HIinter <- as(HIinter,"matrix")
HIgears <- read.csv("HIregion_gears_params.csv", header = TRUE)

# Generate parameters
params <- newMultispeciesParams(HIparams, interaction = HIinter, kappa = 1e12, min_w_pp = 1e-14, w_pp_cutoff = 455400) 
```

Set up time steps:
```{r}
time_steps <- seq(0,85,1)

# I ran into trouble using any time steps that didn't start with 0.
```

Set up the resource:
```{r}
# Load resource spectra
LENS_plankton <- load(here("ClimateForcing", "Plankton", "DataFiles",
                           "LENS_resource_spectra_S1.075I1.975.Rdata"))

# We'll just do the first ensemble member for now.  
# Once that looks good, we'll test it with the other nine members.
LENS_plankton_1 <- out_LENS[,,1]
LENS_plankton_1 <- as(LENS_plankton_1, "matrix")
sizes <- names(params@initial_n_pp)
n_pp_array <- array(NA, dim = c(length(time_steps), length(sizes)), dimnames = list(time = time_steps, w = sizes))

# Fill array
# Need an extra time step preceding the simulation
n_pp_array[1,] <- LENS_plankton_1[1,]
for (t in seq(1,length(time_steps) - 1,1)) {
  n_pp_array[t + 1,] <- LENS_plankton_1[t,]
}
```

Set up temperature-related information:
```{r}
# Load ocean temp array
# Here, too, we'll just use the first ensemble member.

ocean_temp <- read.table(here("ClimateForcing", "Temperature", "DataFiles",
                              "LE2_001_ocean_temp_realm_array.dat"))
ocean_temp <- as(ocean_temp, "matrix")
r <- colnames(ocean_temp)
ocean_temp_array <- array(NA, dim = c(length(time_steps), length(r)), dimnames = list(time = time_steps, realm_names = r))
ocean_temp_array[1,] <- ocean_temp[1,]
for (t in seq(1,length(time_steps) - 1, 1)) {
  ocean_temp_array[t + 1,] <- ocean_temp[t,]
}

# Load vertical migration and exposure
load("HIregion_vertical_migration_array.Rdata") # Note: this is named vertical_migration_array

HIexposure <- read.table("HIregion_exposure_array.dat")
HIexposure <- as(HIexposure, "matrix")
```

Set up the needed fishing selectivity function:
```{r}
# We need to define our double knife-edge selectivity function to match that 
# used in Woodworth-Jefcoats et al. 2019
double_knife_edge <- function(w, knife_edge_size1, knife_edge_size2, ...) { 
	sel <- rep(0, length(w)) 
	
	# Phased in linearly
	F0 <- which(w < knife_edge_size1) # to find one size smaller than that fished, for the 0 value 
	F1 <- which(w < knife_edge_size2) # to find end of escalation size range 
	lo_sel <- max(F0):max(F1) 
	sel[lo_sel] <- seq(0, 1, length = length(lo_sel)) # linear increase from 0 to F 
	sel[w >= knife_edge_size2] <- 1 
	return(sel) 
} 
```

Add gear parameters:
```{r}
gear_params(params) <- HIgears
```

Add temperature and resource information to the parameters:
```{r}
params <- upgradeTherParams(params, ocean_temp_array = ocean_temp_array,
                            n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = HIexposure, 
                            aerobic_effect = TRUE, metabolism_effect = TRUE)

# Nice test to see that things are working correctly
plotTherPerformance(params)
plotTherScalar(params)
```

Project to steady-state, with constant fishing at F = 0.2 (the estimated F for most of the species modeled, per stock assessments):
```{r}
# Project to steady
params_steady_F <- projectToSteady(params, t_max = 600, dt = 1/12, tol = 1e-7, effort = 0.2) 
sim_steady_F <- projectToSteady(params, t_max = 600, dt = 1/12, tol = 1e-7, effort = 0.2, return_sim = TRUE) 

# Test to see how things look
plot(sim_steady_F)
```

Run simulations:
```{r}
# And do a simulation using the steady parameters
projection_steady_F <- project(params_steady_F, dt = 1/12, t_max = 86, t_save = 1/12, effort = 0.2)

# Test to see how things look
plot(projection_steady_F)
```
